---
title: "Tidymodels and Classification on Palmer Penguin Data"
author: "Gaurav Sharma"
date: "27/08/2020"
output: github_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	fig.height = 5,
	fig.width = 8,
	message = FALSE,
	warning = FALSE,
	cache = TRUE,
	dpi = 180
)
library(tidyverse)
library(tidymodels)
```

### Reading the files
```{r}
ttfile <- tidytuesdayR::tt_load('2020-07-28')
penguins <- ttfile$penguins
```

### let's explore the data to see if we can find some good features to classify male and female penguins
```{r Boxplot}
penguins %>% 
    drop_na() %>%
    pivot_longer(bill_length_mm:body_mass_g, names_to = 'spec', values_to = 'value') %>% 
    ggplot(aes(x= value, y = sex, fill = sex)) +
    geom_boxplot(color = "gray") +
    facet_wrap(~spec, scales = "free_y") +
    coord_flip() +
    labs(title = "Body Specs clearly show a difference in male and female penguins")
```

```{r Dotplot}
penguins %>% 
    filter(!is.na(sex)) %>% 
    ggplot(aes(flipper_length_mm, bill_length_mm, color = sex, size = body_mass_g)) +
    geom_point(alpha = 0.5) +
    facet_wrap(~species) +
    labs(title = "Flipper and bill length can be used to differentiate between male and female penguins")
```

### Lets drop the NAs and finalize the data that we are going to use for modelling
```{r}
penguins_df <- penguins %>% 
    drop_na() %>% 
    select(-c(year, island)) %>% 
    mutate(sex = as_factor(sex))
```

### Lets create a split to get our training and testing datasets
```{r}
set.seed(123)
penguin_split <- penguins_df %>% 
    initial_split(strata = sex)

penguin_train <- training(penguin_split)
penguin_test <- testing(penguin_split)
penguin_split
```

### As we don't have a lot of data hence it makes sense to create resamples to evaluatethe model properly. We will save the test set for final generalization.
```{r}
set.seed(234)
penguin_boot <- penguin_train %>% 
    bootstraps(times = 100)
penguin_boot
```

### lets create classification models now. We will create two models. A logistic regression model and a random forest model. Let's create specs for these two models.


```{r}
glm_spec <- logistic_reg() %>% 
    set_engine(engine = 'glm') %>% 
    set_mode(mode = "classification")
glm_spec

rf_spec <- rand_forest() %>% 
    set_mode(mode = "classification") %>% 
    set_engine(engine = "ranger")
rf_spec
```

### Lets create a workflow now to combine the specs and data and define a standardized workflow for modelling
```{r}
penguin_wf <- workflow() %>% 
    add_formula(sex~.)
penguin_wf
```

### Lets train our first model which is a logistic regression model
```{r}
glm_rs <- penguin_wf %>% 
    add_model(glm_spec) %>% 
    fit_resamples(resamples = penguin_boot,
                  control = control_resamples(save_pred = TRUE), 
                  verbose = TRUE)

rf_rs <- penguin_wf %>% 
    add_model(rf_spec) %>% 
    fit_resamples(resamples = penguin_boot,
                  control = control_resamples(save_pred = TRUE), 
                  verbose = TRUE)

```

### Lets have a look at the performances of these models
```{r}
glm_rs %>% 
    collect_metrics()
```

```{r}
rf_rs %>% 
    collect_metrics()
```

```{r roc for resamples}
glm_rs %>% 
    collect_predictions() %>% 
    group_by(id) %>% 
    roc_curve(sex, .pred_female) %>% 
    ggplot(aes(1-specificity, sensitivity, color = id)) +
    geom_abline(lty=2, size = 1, color = "gray") +
    geom_path(show.legend = F, size = 1, alpha = 0.6)
    
```

### Finally lets fit this model to our test data and see how performs
```{r}
penguin_final <- penguin_wf %>% 
  add_model(glm_spec) %>% 
  last_fit(penguin_split)
  
penguin_final
```

```{r}
penguin_final %>% 
  collect_metrics()
```

```{r}
penguin_final %>% 
  collect_predictions() %>% 
  conf_mat(sex, .pred_class)
```

```{r roc for test data}
penguin_final %>% 
  collect_predictions() %>% 
  roc_curve(sex, .pred_female) %>% 
  autoplot()
```

